Investigations on Limited Detection Schemes for Light Scattering of Optically Trapped Asymmetric Particles

### 2022-11-14 Manuscript.tex
Add files via upload

\title{Investigating the Applications and Limits of Single Asymmetric Particle Light Scattering using Limited Detection Schemes}

@@

\cortext[cor1]{Corresponding author}
\affiliation[aff1]{organization={Department of Chemical Engineering,
			University of Strathclyde},
            addressline={75 Montrose Street}, 
            city={Glasgow},
            postcode={G1 1XL}, 
            country={Scotland}}

\begin{abstract}
\justifying
While optical trapping is a well understood method for force transduction and detection, \textcolor{blue}{ further} characterisation of trapped entities \textcolor{blue}{with complex structure and dynamics} \textcolor{blue}{ would provide valuable information for example on aggregation processes, microscale biological processes and so on. Such characterisation using light scattering from the trapped object} poses a two-fold challenge - one experimental, concerning the arrangement of scattered light detectors, and the other theoretical, involving solution of the inverse light scattering problem. \textcolor{blue} {Experimentally,} combining static light scattering techniques with optical trapping poses significant engineering challenges due to the space constraints in a conventional optical trapping setup;  \textcolor{blue} {while the theoretical inverse scattering problem is a non-trivial task given the potentially complex structure of the entity being characterised. To address these issues} we propose here a plausible scenario of detecting scattered light from an optically trapped asymmetric microstructure using a novel, multi-angle, optical-fibre based detection scheme, \textcolor{blue}{combined with}  a Bayesian inference-based analysis method demonstrating how data simulated to mimic light scattering signals in such a scenario may be used to solve the inverse light scattering problem. \textcolor{blue}{As an example problem} we discuss the application of our method to infer the instantaneous orientation of a simulated trapped asymmetric microsphere dimer from the scattered light signal. We argue that this method can be extended for determining in principle any characteristics of the trapped microstructure that influence the light scattering pattern.
\end{abstract}

\begin{highlights}
\item Asymmetric dimers undergo a full inversion in the presence of an potential well. This only occurs if the size ratio is greater than 1:2.  
\item Orientation can be classified by applying Bayesian Statistics to light intensity readings. 
\item Choice of detection angles sets a lower limit on possible signal error. 
\item Future applications in continuos processes where dimer shape is changing with time.   
\end{highlights}

\begin{keyword}
	Optical Trapping \sep Light Scattering \sep Measurements \sep Bayesian Statistics \sep Theoretical 
\end{keyword}

@@

Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to those that test the fundamental assumptions of Quantum Mechanics \cite{yin2013large} thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few piconewtons. The ability to characterise the trapped entity has provided useful insights in several fields of research; such as metrological measurements \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical}. Chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering \cite{gupta2014raman}, while dynamical characterisation has been demonstrated using a Quadrant Photo Detector,  by following the centre-of-mass Brownian motion of the trapped entity  \cite{friedrich2012tuning} and measuring rotation  of the centre-of-mass \cite{yifat2021facile}.

@@

To demonstrate the analysis, we study a simple paradigmatic example of a trapped entity with more complex dynamics and scattering than a single sphere, i.e. a trapped asymmetric dimer; and we explore how to optimise the experimental measurement to obtain the best estimate of the dimer's instantaneous orientation, as a further paradigmatic example of interpretation of the scattering data to obtain dynamical information on the trapped entity. There is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in \cite{raudsepp2022estimating} where imaging was employed to study the orientation of trapped dimers and a resolution of () was achieved. Here, we simulate the Brownian motion of a dimer and calculate the resulting scattered light, and show how Bayesian inference can be used to optimise extraction of the true dimer orientation form the light scattering signals given a choice of scattering measurement angles.

\textcolor{blue} {not very happy with the above but I am trying to set out clearly in advance the narrative of what we are doing, starting with the general problem, moving to the specific example we are studying, and introducing how we then go about it. we need to emphasise how oreitnation of a dimer is our choice of example problem, where a dimer is a paradigm of a 'complex object' and its orientation is a paradigm of 'data' that in some way describes it}

@@

@@

\subsection{Impact of signal error on model predictions}
\label{sec:3.2}
Signal error is an important parameter to control when it comes to understanding how well our model can predict the dimers motion. To this end we ran our model while varying the value of $\epsilon$. \figurename{ 3} shows the model's performance with a selection of angles $15^{\circ},\ 55^{\circ}, \ \& \ 90^{\circ}$ and $\beta$ set at 1. The dotted line indicates the minimum radian distance ($0.137$ radians) between two neighbouring reference orientations, if we are under this line then we know our model is just off from selecting the correct orientation

@@

As can be seen from Figure~\ref{fig:epsilon} there is seemingly a sweet spot for the signal error; with a $5\%$ error giving an inaccurate estimation, and with a $25\%$ error giving an alright estimation but with sudden shifts away from the ideal result. The reason why our low error estimation performed so poorly is that the minimum difference between our values of $y_k$ is larger than our signal error. As such the model could not distinguish between different reference orientations and was essentially just picking options at random. This minimum difference determined by our choice of angles and sets a lower limit of how accurate our signals can be.  In the future we hope to look at increasing the number of reference orientations to see what the true lower bound is.

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.4\textwidth]{./Images/corneranglesfreed-1.png}
	\caption{Corner plot mapping combinations of $\theta$ that generate the best improvement}
	\label{fig:corner}
\end{figure}

@@

\section{Conclusion}
\label{4}
We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern. In theory the model can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity. The MSTM package is flexible in calculating the light scattering for a variety of micro particles; so long as there is a distinguishable difference in the light scattering pattern we can apply this method to characterise the particle to some degree of accuracy. This has potential to be applied to particles that are changing while in the optical trap. We hope to apply this work in the future to  characterise processes such as nucleation within an optical trap.\\

>>>>

\title{Investigations on Limited Detection Schemes for Light Scattering of Optically Trapped Asymmetric Particles}

@@

\cortext[cor1]{Corresponding author}
\affiliation[aff1]{organization={Department of Chemical Engineering,
			University of Strathclyde},
            addressline={75 Montrose Street}, 
            city={Glasgow},
            postcode={G1 1XJ}, 
            country={Scotland}}

\begin{abstract}
	\label{sec:Abstract}
\justifying
While optical trapping is a well understood method for force transduction and detection, further characterisation of trapped entities with complex structure and dynamics would provide valuable information for example on aggregation processes, micro-scale biological processes and so on. Such characterisation using light scattering from the trapped object poses a two-fold challenge - one experimental, concerning the arrangement of scattered light detectors, and the other theoretical, involving solution of the inverse light scattering problem. Experimentally, combining static light scattering techniques with optical trapping poses significant engineering challenges due to the space constraints in a conventional optical trapping setup;  while the theoretical inverse scattering problem is a non-trivial task given the potentially complex structure of the entity being characterised. To address these issues we propose here a plausible scenario of detecting scattered light from an optically trapped asymmetric microstructure using a novel, multi-angle, optical-fibre based detection scheme, combined with a Bayesian inference-based analysis method demonstrating how data simulated to mimic light scattering signals in such a scenario may be used to solve the inverse light scattering problem. As an example problem we discuss the application of our method to infer the instantaneous orientation of a simulated trapped asymmetric microsphere dimer from the scattered light signal. We argue that this method can be extended for determining in principle any characteristics of the trapped microstructure that influence the light scattering pattern.
\end{abstract}

\begin{highlights}
\item Asymmetric dimers undergo a full inversion in the presence of an potential well. This only occurs if the size ratio is greater than 1:2.  
\item Orientation can be classified by applying Bayesian Statistics to light intensity readings. 
\item Choice of detection angles sets a lower limit on possible signal error. 
\item Method can be reworked to determine other characteristics of trapped entities. 
\end{highlights}

\begin{keyword}
	Optical Trapping \sep Light Scattering \sep Measurements \sep Bayesian Statistics 
\end{keyword}

@@

Since their invention in the late 1980's, Optical Tweezers have found application in experiments ranging from single molecule biophysics \cite{Bustamante2021Biophysics} to those that test the fundamental assumptions of Quantum Mechanics \cite{yin2013large} thanks mainly to the ability of the tweezer to transduce and detect forces on the order of a few pico newtons. The ability to characterise the trapped entity has provided useful insights in several fields of research; such as metrological measurements \cite{arita2020coherent} and colloidal aggregation \cite{burns1990optical}. Chemical characterisation of the trapped entities has been achieved using spectroscopic techniques such as Raman Scattering \cite{gupta2014raman}, while dynamical characterisation has been demonstrated using a Quadrant Photo Detector,  by following the centre-of-mass Brownian motion of the trapped entity  \cite{friedrich2012tuning} and measuring rotation  of the centre-of-mass \cite{yifat2021facile}.

@@

To demonstrate the analysis, we study a simple paradigmatic example of a trapped entity with more complex dynamics and scattering than a single sphere, i.e. a trapped asymmetric dimer; and we explore how to optimise the experimental measurement to obtain the best estimate of the dimer's instantaneous orientation, as a further paradigmatic example of interpretation of the scattering data to obtain dynamical information on the trapped entity. There is a paucity of literature on measuring the orientation of trapped non-spherical particles till the recent attempt in \cite{raudsepp2022estimating} where imaging was employed to study the orientation of trapped dimers. Here, we simulate the Brownian motion of an asymmetric dimer, calculate its light scattering pattern, and show how Bayesian inference can be used to optimise extraction of the true dimer orientation from the light scattering signals given a choice of scattering measurement angles.

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.9\textwidth]{./Images/Dimer.png}
	\caption{Example dimer in orientation $\hat{s}$}
	\label{fig:dimer}
\end{figure}

@@

\subsection{Impact of signal error on model predictions}
\label{sec:3.2}
A key assumption of our Bayesian model is that we assume our detected light signal will inherently have some degree of noise. This is the central premise that allows us to form equation~\eqref{eq:conditional}, understanding the limits of acceptable signal error is necessary for implementing this to experimental work.
To this end we ran our model while varying the value of $\epsilon$; \figurename{ 3} shows the model's performance with a selection of angles $15^{\circ},\ 55^{\circ}, \ \& \ 90^{\circ}$ and $\beta$ set at 1. The dotted line indicates the minimum radian distance ($0.137$ radians) between two neighbouring reference orientations, if we are under this line then we know the instantaneous orientation $\hat{s}$ is between two reference orientation.

@@

As can be seen from Figure~\ref{fig:epsilon} there is an optimal point for the signal error; with a $5\%$ error giving an inaccurate estimation, and with a $25\%$ error giving an good estimation but with periodic spikes that indicate a very poor classification of our dimer's orientation. The reason why our low error estimation performed so poorly is that the minimum difference between our reference signals $y_k(\hat{n})$ is larger than our signal error.  As such the model could not distinguish between different reference orientations and was essentially just picking options at random. For our particular choice of angles the minimum difference was around 0.1734 which works out to around a $15\%$ signal error, lowering the signal error any more than this results in the model's performance dropping substantially. This minimum difference is predetermined by our choice of angles and sets a lower limit of how accurate our signals can be. In the future we hope to look at increasing the number of reference orientations to see if this could allow us to achieve a higher degree of accuracy in our estimation of the instantaneous orientation.

@@

\begin{figure}[h]
	\centering
	\includegraphics[width=0.4\textwidth]{./Images/corneranglesfreed-1.png}
	\caption{Corner plot mapping combinations of $\theta$ that generate the best improvement. Contour ranges from $F(K_l) = 1.7 \ \text{to} \ F(K_l)  =  0.1$ }
	\label{fig:corner}
\end{figure}

@@

\section{Conclusion}
\label{4}
We have developed a model for interpreting the dynamics of a trapped particle based purely on the light scattering pattern. In theory the model can be applied to any characteristic that impacts the light scattering pattern produced by the trapped entity. The MSTM package is flexible in calculating the light scattering for a variety of micro-particles; so long as there is a distinguishable difference in the light scattering pattern we can apply this method to characterise the particle to some degree of accuracy. This has potential to be applied to characterise particle size and shape that are changing while in the optical trap. The authors would like to acknowledge and thank the support for this research from the funding provided by the Leverhulme Trust
